---
title: "week_5_lab"
author: "Lindsey Greenhill"
date: "3/25/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(rdrobust)
library(sandwich)
library(lmtest)

data <- read_dta("probation.dta")
```

## Question 1

We don't waant to compare all students who are on probation with all students who arae not on probation because the students at the top end of the gpa theshold are different in many aspects than students at the bottom end of the gpa theshold. As such, the underlying assumption that the control and treatment groups are comparable would not hold. 

## Question 2

The running variable is gpa average

## Question 3

### Part a

The graphs below check the smoothness of fraction graduating in 4 years, male,
and high school grade precentage. There doesn't appear to be a large break in
any of these variables at the GPA threshold, meaning that it doesn't seem like
any of these factors are driving the differences in the groups.

```{r echo=FALSE}

data$dist_from_cut <- data$GPA - 1.6

#totcredits_year1, male, and hsgrade_pct

dat_narrow <- data %>%
  filter(dist_from_cut <=1.2 & dist_from_cut >= -1.2)

rdplot(dat_narrow$male, #outcome variable
       dat_narrow$dist_from_cut, #running variable
        p = 1, #p = 1 is linear best fit line. p >= 2 is nonlinear
        nbins = c(20, 20), #number of bins on each side of threshold
        binselect = "es", #option to use "equal spaced" binning
        y.lim = c(0, 0.6), #Set y-axis scale
        x.label = "Grade Point Average minus 1.6", 
        y.label = "Male",
       title = "Male"
       )

rdplot(dat_narrow$hsgrade_pct, #outcome variable
       dat_narrow$dist_from_cut, #running variable
        p = 1, #p = 1 is linear best fit line. p >= 2 is nonlinear
        nbins = c(20, 20), #number of bins on each side of threshold
        binselect = "es", #option to use "equal spaced" binning
        x.label = "Grade Point Average minus 1.6", 
        y.label = "High School Grade Percentage",
       title = "High School Grade Percentage"
       )


```

### Part b

There doesn't appear to be a spike in the density of observations just above
the GPA theshold. 

```{r echo=FALSE}

# creating histogram

dat_narrow %>%
  ggplot(aes(x = dist_from_cut,
             y = ..density..)) +
  geom_histogram(bins = 200) +
  labs(title = "Distribution of dist_from_cut",
       subtitle = "There doesn't appear to be a spike in density
       just above the cutoff") +
  theme_classic()


```


## Question 4

```{r echo=FALSE}

rdplot(dat_narrow$gradin4, #outcome variable
       dat_narrow$dist_from_cut, #running variable
        p = 1, #p = 1 is linear best fit line. p >= 2 is nonlinear
        nbins = c(20, 20), #number of bins on each side of threshold
        binselect = "es", #option to use "equal spaced" binning
        y.lim = c(0, 0.6), #Set y-axis scale
        x.label = "Grade Point Average minus 1.6", 
        y.label = "Fraction Graduating in 4 years",
       title = "Fraction Graduating in 4 years"
       )

```

## Question 5

```{r echo=FALSE}

# filtering the data

data_left <- data %>%
  filter(GPA < 1.6, GPA >= .4)

## part a

# regression

mod_1 <- lm(gradin4 ~ GPA, data = data_left)

answer_a <- -.1327 + (1.6*.2333)

## part b

data_right <- data %>%
  filter(GPA >= 1.6, GPA <= 2.8)

mod_2 <- lm(gradin4 ~ GPA, data = data_right)

answer_b <- -.05917 + (1.6*.21278)

## part c

answer_c <- answer_b - answer_a

```

a) `r answer_a`

b) `r answer_b`

c) The difference between b and a, or the predicted effect = `r answer_c`

## Question 6

```{r echo=FALSE}

#To estimate the discontinuity requires generating some additional variables
#generate indicator for being above probation threshold
data$T <- 0
data$T[which(data$dist_from_cut >= 0)] <- 1

#Generate interaction term for linear
data$interaction <- data$dist_from_cut*data$T

##Subset data to [-1.2,1.2] with new variables added
dat_narrow <- subset(data,dist_from_cut<=1.2 & dist_from_cut>=-1.2)

#Estimate regression
linear <- lm(gradin4 ~ T + dist_from_cut + interaction , data = dat_narrow)
coeftest(linear, vcov = vcovHC(linear, type="HC1"))
```

The coefficient for T equals the answer in 5c. 

## Question 7

The standard error for the coefficient = .013. Because .0407 +/- 1.96*.013 does
not include 0, we can conclude that the effect is statistically significant at
the 95% confidence level.

## Question 8

Because the effect of the T coeffient is positive and statistically significant,
we can see that kids who received probation are graduating at a slower rate than
kids who did not receive probation. As such, we can conclude the school's
probation policy was effective in a sense that is punished kids below the
threshold. However, it was not effective at improving their academic outcomes
when measured with graduation rates. With this being said, the effect (.04) is
very small, so while we can claim that it is statistically significant, in my
opinion, it does not seem substantively significant and is certainly not
deterministic of any one student's academic outcomes.




