---
title: "lab_9"
author: "Lindsey Greenhill"
date: "4/21/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(haven)
library(statar)
library(lmtest)
library(sandwich)
library(Synth)
library(stargazer)
library(gt)
library(SCtools)
cig <- read_dta("ec50_smoking.dta")

```

## Question 1

I look at Alabama in this graph. There were tax hikes in 1981, 2004, and 2016.
The largest of these hikes occured in 2004, when the tax rose from .165 in 2003
to .425 in 2004. The hike in 2015 was almost as large: the tax from from .425 in
2015 to .675 in 2016.

```{r echo=FALSE}
# I am looking at Alabama

cig %>%
  filter(state_fips == 1,
         year >=1970, year <= 2018) %>%
  ggplot(aes(x = year, y = state_tax_dollars)) +
  geom_point() +
  geom_line() +
  labs(title = "AL State Cigarette Tax Per Pack",
       subtitle = "1970 - 2018",
       y = "state cigarette tax per pack") +
  theme_classic()

# from this graph, it seems like there hve been many tax policy changes in CT
# since 2000. It looks like the largest of these jumps happened from 209 to
# 2010, when the state tax rose from $2 to $3
```

## Question 2

### Part a

```{r echo=FALSE}

cig$al <- 0
cig$al[which(cig$state=="AL")] <- 1
cig_narrow <- subset(cig, year>=2006 & year <= 2018)

#Replicate figure 1

ggplot(cig_narrow, aes(x=year,y=pack_sales,
                       shape= factor(al, labels = c("Rest of U.S.", "Alabama")))) +
  geom_vline(xintercept=2016) +
  stat_summary(fun = "mean",geom="point") +
  stat_summary(fun = "mean",geom="line")  +
  labs(x = "Year", y = "Cigarette Consumption", shape = "") +
  theme(legend.position="bottom") +
  labs(title = "Alabama vs. Rest of US") +
  theme_classic()
```

### Part b

```{r echo=FALSE}

#Compare Alabama to only neighboring states
# GA, FL, TN, and MS
cig_narrow <- subset(cig_narrow, state == "AL" 
                     | state == "GA" 
                     | state == "FL" 
                     | state == "MS" 
                     | state == "TN")


#Replicate figure 2
ggplot(cig_narrow, 
       aes(x=year,y=pack_sales, 
           shape = factor(al, labels = c("Surrounding States", "Alabama")))) +
  geom_vline(xintercept=2016) +
  stat_summary(fun = "mean",geom="point") +
  stat_summary(fun = "mean",geom="line")  +
  labs(x = "Year", y = "Cigarette Consumption", shape = "") +
  theme(legend.position="bottom") +
  labs(title = "Alabama vs. Neighboring States") +
  theme_classic()

```

### Part c

The parallel trends assumption looks plausible for part a, but the the trends in
the surrounding states look like they have some strange spikes that Alabama
doesn't have.

## Question 3

Based on the previous question, I am going to use the entirety of the US as my
control group. See the table below for the statistics on the effect of the 2016
tax hike in Alabama vs. the Rest of the U.S.. Using the difference in
differences calculation, I calculate that the impact of the ccigarette tax
policy change = 6.88 - 5.77 = 1.11. In context, this implies that the 2016
policy results in a decrease of 1.11 in consumption in Alabama.

```{r echo=FALSE}

# calculating stats

q3_prev_5 <- cig %>%
  filter(year >=2011, year < 2016)
q3_after <- cig %>%
  filter(year >= 2016)

stat_1 <- q3_prev_5 %>% 
  filter(state == "AL") %>%
  summarise(mean_cig_consumption = mean(pack_sales, na.rm = T)) %>%
  as.numeric()

stat_2 <- q3_after %>% 
  filter(state == "AL") %>%
  summarise(mean_cig_consumption = mean(pack_sales, na.rm = T)) %>%
  as.numeric()

stat_3 <- q3_prev_5 %>% 
  filter(state != "AL") %>%
  summarise(mean_cig_consumption = mean(pack_sales, na.rm = T)) %>%
  as.numeric()

stat_4 <- q3_after %>% 
  filter(state != "AL") %>%
  summarise(mean_cig_consumption = mean(pack_sales, na.rm = T)) %>%
  as.numeric()

# creating new table

q3_table <- tibble(group = c("Alabama", "Rest of U.S."),
                   mean_consumption_before_2016 = c(stat_1, stat_3),
                   mean_consumption_after_2016 = c(stat_2, stat_4),
                   difference = mean_consumption_before_2016 - mean_consumption_after_2016)

q3_table %>%
  gt()
```
## Question 4

The coefficient for dd (11.11) is the same as what I calculated above. 

\newpage

```{r echo=FALSE, results = "asis"}

q4_df <- cig %>%
  filter(year >=2011)

q4_df$post <- 0
q4_df$post[which(q4_df$year >= 2016)] <- 1

#Generate interaction term
q4_df$dd <- q4_df$al*q4_df$post

#Subset to years you want
#cig_narrow2 only includes the years right around the policy change
q4_df_2 <- subset(q4_df, year>= 2015 & year <= 2017)

#Estimate simple differences in differences
reg1 <- lm(pack_sales ~ post + al + dd, data=q4_df)

stargazer(reg1,
          type = "latex")

#Report heteroskedasticity robust standard errors
coeftest(reg1, vcov = vcovHC(reg1, type="HC1"))
```

\newpage

## Question 5

The difference in difference estimate is -1.11. This implies that Alabama's tax
policy resulted on average in a 1.11 unit decrease in cigarette consumption,
holding all else constant.




```{r echo=FALSE}
#What if we use more years? cig_narrow includes the 1987-2000
#Estimate difference in difference regression extended to multiple control states and many years
#factor(year) generates a fixed effect for each year, excluding one as the reference level

#factor(state) generates a fixed effect for each state, excluding one as the reference level

reg2 <- lm(pack_sales ~ dd + factor(year) + factor(state_fips), 
           data=q4_df)
coeftest(reg2, vcov = vcovHC(reg2, type="HC1"))
```

## Question 6

GA, MO, and WV receive the most weight. 

With regards to the graphs, there seems to be at least a growing of the gap when the 2016 
tax hike takes place. It is easier to see this gap in this graph than in the graphs from 
question 2. 

The pvalue = .05, so it could be considered statistically significant. 

\newpage


```{r echo=FALSE}

#Data prep
#Natural log of state income per capita
cig$lnincome <- log(cig$personal_income)


#Restrict to CA and 38 states that did not have other policy changes
policy <- subset(cig, sample == 1)



#-------------------------------------------------------------------------------
# Graphical analyses
#-------------------------------------------------------------------------------


#Subset data to particular time range
q6 <- subset(policy, year >= 2006)

#Replicate figure 1
ggplot(q6, aes(x=year,y=pack_sales, 
                       shape= factor(al, labels = c("Rest of U.S.", "Alabama")))) +
  geom_vline(xintercept=2016) +
  stat_summary(fun.y = "mean",geom="point") +
  stat_summary(fun.y = "mean",geom="line")  +
  labs(x = "Year", y = "Cigarette Consumption", shape = "",
       title = "Alabama vs. Rest of U.S. 2006-2018") +
  theme(legend.position="bottom") +
  theme_classic()



#-------------------------------------------------------------------------------
# Synthetic control: Estimation
#-------------------------------------------------------------------------------

#This example follows: 

#https://www.rdocumentation.org/packages/Synth/versions/1.1-5/topics/synth
#https://www.rdocumentation.org/packages/Synth/versions/1.1-5/topics/dataprep


#https://rpubs.com/danilofreire/synth


#Store FIPS code for treated state
treated_state <- 1 #This is California in the example

# Make a list of control states' fips codes, excluding treated state

#Start with a list of all unique FIPS codes
list_state <- unique(policy$state_fips)

#Now remove the treated state's FIPS code from the list
list_state <- list_state[list_state != treated_state]

# I'm using wine, age14to24, and cost per pack 
# create matrices from panel data that provide inputs for synth()
policy <- as.data.frame(policy)

dataprep.out<-
  dataprep(
    foo = policy,
    predictors = c("wine", "age15to24", "cost_per_pack"),
    predictors.op = "mean",
    dependent = "pack_sales",
    unit.variable = "state_fips",
    time.variable = "year",
    special.predictors = list(
      list("pack_sales", 2009, "mean"),
      list("pack_sales", 2014, "mean"),
      list("pack_sales", 2016, "mean")
    ),
    treatment.identifier = treated_state,
    controls.identifier = list_state,
    time.predictors.prior = c(2006:2016),
    time.optimize.ssr = c(2006:2016),
    unit.names.variable = "state",
    time.plot = 2006:2018
  )

## run the synth command to identify the weights
## that create the best possible synthetic 
## control unit for the treated.
synth.out <- synth(dataprep.out)

## there are two ways to summarize the results
## we can either access the output from synth.out directly
round(synth.out$solution.w,2)
# contains the unit weights  

## the output from synth opt 
## can be flexibly combined with 
## the output from dataprep to 
## compute other quantities of interest
## for example, the period by period 
## discrepancies between the 
## treated unit and its synthetic control unit
## can be computed by typing
gaps<- dataprep.out$Y1plot-(
  dataprep.out$Y0plot%*%synth.out$solution.w
) ; gaps

## also there are three convenience functions to summarize results.
## to get summary tables for all information 
## (V and W weights plus balance btw. 
## treated and synthetic control) use the 
## synth.tab() command
synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out)
# print(synth.tables)

## to get summary plots for outcome trajectories 
## of the treated and the synthetic control unit use the 
## path.plot() and the gaps.plot() commands

## plot in levels (treated and synthetic)
path.plot(dataprep.res = dataprep.out,synth.res = synth.out, tr.intake = 1989)
#dev.copy(png,'prop99_graph1.png')
#dev.off()

## plot the gaps (treated - synthetic)
gaps.plot(dataprep.res = dataprep.out,synth.res = synth.out, tr.intake = 1989)
#dev.copy(png,'prop99_graph2.png')
#dev.off()

tdf <- generate.placebos(dataprep.out,synth.out, Sigf.ipop = 2)

## Plot the gaps in outcome values over time of each unit --
## treated and placebos -- to their synthetic controls
p <- plot_placebos(tdf,discard.extreme=TRUE, mspe.limit=10, xlab='Year')
p


## Test how extreme was the observed treatment effect given the placebos:
ratio <- mspe.test(tdf)
ratio$p.val
mspe.plot(tdf, discard.extreme = FALSE)

```


