---
title: "week_3_lab"
author: "Lindsey Greenhill"
date: "2/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(skimr)
library(stargazer)

mto <- read_dta("mto.dta")
```

## Question 1 (work on this)

According to lecture, individuals who were young children in 1994-1998 and moved 
as a result of MTO saw improved economic outcomes. These effects did not occur for 
older childer in adults who moved. 

## Question 2

```{r echo=FALSE}

# it looks like nobody in the control group moved

c_moved <- mto %>%
  filter(voucher == 0) %>%
  summarise(frac_moved = mean(moved, na.rm = T))
```

## Question 3

```{r echo=FALSE}

# 43.06% moved

t_moved <- mto %>%
  filter(voucher == 1) %>%
  summarise(frac_moved = mean(moved, na.rm = T))
```

43.06% of the individuals in the experimental group moved.

## Question 4

```{r echo=FALSE, results = "asis"}
mod_4 <- lm(moved ~ voucher, data = mto)
stargazer(mod_4, type = "latex")
```

- The estimated coefficient for voucher is the intent to treat effect of
treatment. This coefficient is equal to the difference in fraction treated
across treatment and control groups (or the difference between the value we
calculated in question 3 and question 2).
 
 ## Question 5
 
 - Because the coefficient on voucher is less than 1, we know there is
 non-compliance. DO WE NEED TO SAY EXACT COMPLIANCE RATE?
 
 -The MTO experiment had one-sided non-compliance.
 
 ## Question 6
 
```{r echo=FALSE, results= "asis"}
# creating model

mod_6 <- lm(kessler ~ voucher, data = mto)

# displaying model

stargazer(mod_6, type = "latex")
```
 
 The intent to treat effect of the voucher on the kessler index = -.588
 
 ## Question 7
 
```{r echo=FALSE}

# the treatment-on-the-treated effect = itt outcome / itt treatment or question
# 6 / question 4

tot <- -.588 / .4306
```
 
 The treatment-on-the-treated efect of actually using the experimental
 voucher to move = -1.365. I calculated this by dividing the ITT of the outcome
 by the ITT of the treatment. You need to make this adjustment because just 
 using the ITT of the outcome would give a causal effect that is too small
 because it doesn't take the compliance rate into account. 
 
 

