---
title: "week_3_lab"
author: "Lindsey Greenhill"
date: "2/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(ggthemes)
library(skimr)
library(stargazer)

mto <- read_dta("mto.dta")
```

## Question 1 (work on this)

According to lecture, individuals who were young children in 1994-1998 and moved
as a result of MTO saw significantly improved economic outcomes that lasted
through their adult life (meaning that their incomes were higher than they would
have been if they hadn't moved). However, these effects did not carry over for
older children who moved (they did not see any significant increase in income as
adults).


## Question 2

```{r echo=FALSE}

# it looks like nobody in the control group moved

c_moved <- mto %>%
  filter(voucher == 0) %>%
  summarise(frac_moved = mean(moved, na.rm = T))
```

Nobody in the control group moved. 

## Question 3

```{r echo=FALSE}

# 43.06% moved

t_moved <- mto %>%
  filter(voucher == 1) %>%
  summarise(frac_moved = mean(moved, na.rm = T))
```

43.06% of the individuals in the experimental group moved.

## Question 4

```{r echo=FALSE, results = "asis"}
mod_4 <- lm(moved ~ voucher, data = mto)
stargazer(mod_4, type = "latex")
```

- The estimated coefficient for voucher is the intent to treat effect of
treatment. This coefficient = .431 and is equal to the difference in fraction treated
across treatment and control groups (or the difference between the value we
calculated in question 3 and question 2). It is also the compliance rate. 
 
 ## Question 5
 
 - Because the coefficient on voucher is less than 1, we know there is
 non-compliance. We also know from our regression that the compliance rate =
 43.1%
 
 -The MTO experiment had one-sided non-compliance because there is 
 non-compliance in the treatment group but full compliance in the control
 group. 
 
 ## Question 6
 
```{r echo=FALSE, results= "asis"}
# creating model

mod_6 <- lm(kessler ~ voucher, data = mto)

# displaying model

stargazer(mod_6, type = "latex")
```
 
 The intent to treat effect of the voucher on the kessler index = -.588
 
 ## Question 7
 
```{r echo=FALSE}

# the treatment-on-the-treated effect = itt outcome / itt treatment or question
# 6 / question 4

tot <- -.588 / .4306
```
 
 The treatment-on-the-treated effect of actually using the experimental
 voucher to move = -1.365. I calculated this by dividing the ITT of the outcome
 by the ITT of the treatment. You need to make this adjustment because just 
 using the ITT of the outcome would give a causal effect that is too small
 because it doesn't take the compliance rate into account. 
 
 
 ## Question 8
 
 
```{r echo=FALSE}
as_treated <- mto %>%
  group_by(moved) %>%
  summarise(avg_kessler = mean(kessler, na.rm = T))

# avg for those who did not move = 6.6778 and avg for those who did move =
# 6.1515 so the as treated = 6.1515 - 6.6778 = -5.263
```
 
 
 The as treated effect = -.5263
 
 
 ## Question 9
 
 
```{r echo = FALSE}
v_pp <- mto %>%
  filter(voucher == 1, moved == 1) %>%
  summarise(avg_kessler = mean(kessler, na.rm = T))

c_pp <- mto %>%
  filter(voucher == 0) %>%
  summarise(avg_kessler = mean(kessler, na.rm = T))

pp <- v_pp - c_pp


```
 
 The per protocol analysis gives a result of -.729
 
## Question 10

- The treatement-on-the-treated effect = -1.365
- The as treated effect = -.5263
- The per protocol effect = -.729

The treatment on the treated effect yields the biggest estimate. 

## Question 11

The per protocol and as treated approaches both face problems of selection
bias and confounding variables because they invalidate the randomness in 
the experiment that lets us claim causality in the first place. In contrast,
the TOT analysis preserves the original randomess of the experiment. 

## Question 12

```{r echo=FALSE, message=FALSE}

mto %>%
  mutate(voucher = as_factor(voucher)) %>%
  group_by(voucher) %>%
  summarise(fraction_moved = mean(moved, na.rm = T)) %>%
  ggplot(aes(x = voucher, y = fraction_moved)) +
  geom_col(fill = "steelblue2") +
  scale_x_discrete(breaks = c("0", "1"),
                   labels = c("Control", "Treatment")) +
  labs(title = "Fraction Moved Treatment and Control",
       y = "Fraction Moved",
       x = "Group") +
  theme_clean()

mto %>%
  mutate(voucher = as_factor(voucher)) %>%
  group_by(voucher) %>%
  summarise(avg_kessler = mean(kessler, na.rm = T)) %>%
  ggplot(aes(x = voucher, y = avg_kessler)) +
  geom_col(fill = "steelblue2") +
  scale_x_discrete(breaks = c("0", "1"),
                   labels = c("Control", "Treatment")) +
  labs(title = "Mean Kessler Treatment and Control",
       subtitle = "Not adjusted for TOT effect",
       y = "Mean Kessler",
       x = "Group") +
  theme_clean()

tibble(group = c("Control", "Treatment"),
       mean_kessler = c(6.88, (6.88 - 1.365))) %>%
  ggplot(aes(x = group, y = mean_kessler)) +
  geom_col(fill = "azure3") +
  labs(title = "Mean Kessler Treatment and Control",
       subtitle = "Adjusted for TOT effect",
       y = "Mean Kessler",
       x = "Group") +
  theme_clean()
  

```

 

