---
title: "week_2_submission"
author: "Lindsey Greenhill"
date: "2/11/2022"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(haven)
library(statar)
library(stargazer)

# reading in data

nlsy <- read_dta("nlsy97.dta")
```

## Question 1

```{r echo = FALSE}
# Part a

# creating new indicator variable for mother going to college

nlsy <- nlsy %>%
  mutate(mother_less_college = if_else(mother_education <= 12, 1, 0))

# calculating fraction for both groups

frac_1 <- nlsy %>%
  group_by(mother_less_college) %>%
  summarise(frac_kid_college = mean(child_college))

# ANSWER: .182 fraction of the sample whose parents did not go to college
# (measured by mother's education) went on to receive a college degree or
# higher.

# Part b 

sd_1 <- nlsy %>%
  group_by(mother_less_college) %>%
  summarise(frac_kid_college = sd(child_college))

# ANSWER: I calculated that 18.2% of the children whose parents had a high
# school education or less went on to receive a college degree or higher. The
# standard deviation of this statistic = .386. With this in mind,.182 seems
# close to .209 because .209  is well within one standard deviation of .182.
```

### Part a and b

18.18% of the children whose parents had a high school degree or less went on to
receive a college degree or higher. The standard deviation os this statistic is 
38.6%. With this is mind, 18.18% seems close to 20.9% because 20.9% is well within 
one standard deviation of 18.2%. 

## Question 2

```{r echo=FALSE}

# function from last week's lab

percentile_rank <- function(variable){ 
  
  #Convert to ranks, taking care of potential missing values 
  r <- ifelse(is.na(variable), NA, rank(variable, ties.method = "average")) 
  
  #Return percentile rank = rank normalized so max is 100 
  100*r/max(r, na.rm = T) 
} 

# part a -- taken from lab 1

nlsy$kid_inc_rank <-with(nlsy, percentile_rank(kid_income)) 
nlsy$parent_inc_rank <-with(nlsy, percentile_rank(parent_inc)) 

```

Ranks generated in code. 

## Question 3

```{r echo = FALSE}
# part a

nlsy %>%
  ggplot(aes(x = parent_inc_rank, y = kid_inc_rank)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Scatter plot of kid_inc_rank vs. parent_inc_rank") +
  theme_classic()


# part b

ggplot(nlsy, aes(x = parent_inc_rank, y = kid_inc_rank)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "binned scatter plot of kid_inc_rank vs. parent_inc_rank") +
  theme_classic()

# ANSWER: I think that the binned scatter plot is more useful in this instance
# because it is much easier to see the relationship between the two variables.
# In the basic scatter plot, many of the points overlap and it is difficult to
# discern any clear pattern. With that being said, the scatterplot also probably
# gives a more clear picture of the variability in the data. 

```

I think that the binned scatter plot is more useful in this instance
because it is much easier to see the relationship between the two variables.
In the basic scatter plot, many of the points overlap and it is difficult to
discern any clear pattern. With that being said, the scatterplot also probably
gives a more clear picture of the variability in the data. 

\newpage

## Question 4

```{r echo=FALSE, results="asis"}
mod_4 <- lm(kid_inc_rank ~ parent_inc_rank, data = nlsy)
stargazer(mod_4,
          type = "latex")

# ANSWER: The intercept of the regression = 31.418. The estimated slope = .372.
# The interpretation of this model is: for every 1 unit increases in the
# parent_inc_rank variable, there is a .372 unit increase in the kid_inc_rank
# variable on average. In simpler words, a 1 point increase in parent income rank
# results in a .372 point increase in the kid income rank on average.
# Additionally, when the parent_income_rank = 0, the kid_income_rank is equal to
# 31.418 on average. 
```

The intercept of the regression = 31.418. The estimated slope = .371.

\newpage


## Question 5

```{r echo=FALSE}
stat_5a <- 31.4183 + (.3728*25)

# part a: 40.74

stat_5b <- (31.4183 + (.3728*100)) - (31.4183 + (.3728*0))

# part b: 37.28. So the sample suggests that the world is more unequal than the
# tax data does

# check this. should it be inclusive?

stat_5c <- nlsy %>%
  filter(parent_inc_rank < 20) %>%
  mutate(move_up = if_else(kid_inc_rank >=80, 1, 0)) %>%
  summarise(prob = mean(move_up))

#ANSWER: 7.37%

# do I need to adjust for inflation here?

stat_5d <- nlsy %>%
  mutate(earned_more_parents = if_else((kid_income/1.4767) > parent_inc, 1, 0)) %>%
  summarize(frac = mean(earned_more_parents))

# part d: 50.9% of children born in the 1980s earned more than their parents
# after adjusting for inflation
```

### Part a

Statistic 1 for the sample = `r stat_5a`. This is less than the value Chetty 
found using the full population data. 

### Part b

Statistic 2 for the sample = `r stat_5b`. This value is higher than the value 
Chetty found using the full population data. 

### Part c

Statistic 3 for the sample = 7.37%. This value is less than the value Chetty 
found using the full population data. 

### Part d

Statistic 4 for the sample = 50.9% (adjusting for inflation). This value is 
slightly higher than the value Chetty calculated. 

\newpage

## Question 6

```{r echo=FALSE, results = "asis"}
# black men stats #

black_men <- nlsy %>%
  filter(female == 0, black == 1)

black_men_mod <- lm(kid_inc_rank ~ parent_inc_rank, data = black_men)

stat_6a_b <- 25.70348 + (.29432*25)

# ANSWER: 33.03

stat_6b_b <- (25.70348 + (.29432*100)) - (25.70348 + (.29432*0))

# ANSWER: 29.43

stat_5c_b <- black_men %>%
  filter(parent_inc_rank < 20) %>%
  mutate(move_up = if_else(kid_inc_rank >=80, 1, 0)) %>%
  summarise(prob = mean(move_up))

# ANSWER: 5.79%

stat_6d_b <- black_men %>%
  mutate(earned_more_parents = if_else((kid_income/1.4767) > parent_inc, 1, 0)) %>%
  summarize(frac = mean(earned_more_parents))

# ANSWER: 48.9

# white men stats #

white_men <- nlsy %>%
  filter(female == 0, white == 1)

white_men_mod <- lm(kid_inc_rank ~ parent_inc_rank, data = white_men)


stat_6a_w <- 40.11022 + (.26692*25)

# ANSWER: 46.783

stat_6b_w <- (40.11022 + (.26692*100)) - (40.11022 + (.26692*0))

# ANSWER: 26.692

stat_5c_w <- white_men %>%
  filter(parent_inc_rank < 20) %>%
  mutate(move_up = if_else(kid_inc_rank >=80, 1, 0)) %>%
  summarise(prob = mean(move_up))

# ANWER: 10.3%  

stat_6d_w <- white_men %>%
  mutate(earned_more_parents = if_else((kid_income/1.4767) > parent_inc, 1, 0)) %>%
  summarize(frac = mean(earned_more_parents))

# ANSWER: 48.3

```

### Black Men Stats

- Statistic 1 = 33.03
- Statistic 2 = 29.43
- Statistic 3 = 5.785%
- Statistic 4 = 48.9%

### White Men Stats

- Statistic 1 = 46.783
- Statistic 2 = 26.692
- Statistic 3 = 10.27%
- Statistic 4 = 48.3%

### Comparison

  Mobility seems higher for White men than Black men when comparing statistics 
  1, 2, and 3. Mobility seems higher for Black men than for White men when comparing
  statistic 4. 
  
  \newpage
  
## Question 7

```{r echo=FALSE}

# using the inclusive boundaries


seven_a <- nlsy %>%
  filter(parent_inc_rank >= 21.5 & parent_inc_rank <= 27.5) %>%
  summarise(avg = mean(kid_inc_rank, na.rm = T))

# the arithmetic mean of kid_inc_rank for children with parent_inc_rank between
# 21.5 and 27.5 (inclusive) = 40.74838. This is pretty much the same value as
# Statistic 1 calculated in 5a (40.7383.
```

### Part a

The arithmetic mean of kid_inc_rank for children with parent_inc_rank between
 21.5 and 27.5 (inclusive) = 40.74838. This is pretty much the same value as
 Statistic 1 calculated in 5a (40.7383).
 
```{r echo=FALSE}
set.seed(31305186)

# random sample of 50

samp <- sample_n(nlsy, 50)

# model 7

mod_7 <- lm(kid_inc_rank ~ parent_inc_rank, data = samp)

seven_b <- 29.178 + (.324*25)

```
 
### Part b

Statistic 1 for my random sample = 37.278

```{r echo=FALSE}
seven_c <- samp %>%
  filter(parent_inc_rank >= 21.5 & parent_inc_rank <= 27.5) %>%
  summarise(avg = mean(kid_inc_rank, na.rm = T))
```

### Part c

The mean of the kid_inc_rank for children with parent_inc_rank between 21.5 and 27.5
(inclusive) = 25.116 for my random sample. 

### Part d

The value I calculated in 7b is closer than the value I calculated in 7c to Chetty's
estimate. 
