---
title: "FC_atlas"
author: "Lindsey Greenhill"
date: "2/25/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggthemes)
library(tidyverse)
library(vtable)
library(statar)
library(gt)
library(ggpubr)
library(haven)

atlas <- read_dta("atlas.dta")

# subsetting to CT, fairfield county, 
fc <- atlas %>%
  filter(state == 9, county == 001)
```

## Question 2

The kfr_pooled_pooled_p25 is the absolute mobility at the 25th percentile
for each census tract. In other words, this statistic measures the average 
income percentile rank for children whose parents were at the 25th percentile
in income. In this data set, the unit is the census tract. Higher values
correspond to higher upward mobility. This statistic is estimated using a linear
model because the data is pretty linear when plotted child income rank vs. 
parent income rank. As such, a linear regression fits the data well. 

\newpage

## Question 3

```{r echo=FALSE, warning=FALSE}

atlas %>%
  ggplot(aes(x = kfr_pooled_pooled_p25,
             y = ..density..)) +
  geom_histogram(binwidth = 1,
                 fill = "azure3") +
  geom_vline(xintercept = mean(atlas$kfr_pooled_pooled_p25, na.rm = T)) +
  labs(title = "Distribution of kfr_pooled_pooled_p25",
       subtitle = "Mean = 42.86") +
  theme_clean()
  

```

- The mean value of kfr_pooled_pooled_p25 = 42.86. This means that on average
kids with parents in the 25th percentile of income distribution have adult
incomes in about the 43rd percentile of the income distribution. The data looks
relatively normally distributed.

\newpage

## Question 4 

```{r echo=FALSE, results = "asis"}
sumtable(atlas,
         vars = c("kfr_pooled_pooled_p25"),
         summ = c("mean(x)",
                  "sd(x)",
                  "min(x)",
                  "max(x)"),
         out = "latex")
```


- The variable can be negative or above 100 because the linear
model does not have an upper and lower bound to it. That is,
it will not correct for values that are over or under the 
possible outcomes of 0 to 100. For this reason, another model
(maybe a glm model) could be better to use to bound the values to 0 and 100. 

## Question 5

```{r echo=FALSE}
gw_mean <- atlas %>%
  filter(state == 9,
         county == 1,
         tract == 11200) %>%
  select(kfr_pooled_pooled_p25)

# the bellehaven mean = 48.5

ct_mean <- atlas %>%
  filter(state == 9) %>%
  summarise(avg = mean(kfr_pooled_pooled_p25,
                       na.rm = T))

# the CT average is 44.8

USA_mean <- atlas %>%
  summarise(avg = mean(kfr_pooled_pooled_p25,
                       na.rm = T))

# the national mean = 42.86

comparative_means <- tibble(place = c("Belle Haven",
                                      "CT",
                                      "USA"),
                            avg_kfr_pooled_pooled_p25 = 
                              c(48.5, 44.8, 42.86)) %>%
  gt()
comparative_means



```


My home census tract is 11200 in Fairfield county in CT. My neighborhood is
called Belle haven. The value for kfr_pooled_pooled_p25 = 48.5, which is higher
than the state average of 44.8 and national average of 42.86. This implies that
kids in Belle Haven have a better chance than the average Connecticution and
average American of climbing the income latter.

## Question 6

```{r echo=FALSE}

fc_mean_sd <- atlas %>%
  filter(state == 9,
         county == 1) %>%
  summarise(avg = sd(kfr_pooled_pooled_p25,
                       na.rm = T))

# the fairfield county mean = 8.15

ct_mean_sd <- atlas %>%
  filter(state == 9) %>%
  summarise(avg = sd(kfr_pooled_pooled_p25,
                       na.rm = T))

# the CT average is 7

USA_mean_sd <- atlas %>%
  summarise(avg = sd(kfr_pooled_pooled_p25,
                       na.rm = T))

# the national mean = 7.12

comparative_sd <- tibble(place = c("Fairfield County",
                                      "CT",
                                      "USA"),
                            sd_kfr_pooled_pooled_p25 = 
                              c(8.15, 7, 7.12)) %>%
  gt()
comparative_sd

```

The standard deviation in fairfield county = 8.15, which is higher than the CT
sd of 7 and national sd of 7.12. This means that there is more variation in
upward mobility in my home county than in CT as a whole and the country as a
whole. Subtantively, this means that while there are areas in my county where
upward mobility is quite high, there are also probably areas in my county where
upward mobility is quite low. In other words, a larger sd is probbaly more
indicative of more values at either end of the spectrum.

\newpage

## Question 7

```{r echo=FALSE}

# looking at redlining 

A <- atlas %>%
  filter(HOLC_A > .5,
         !is.na(HOLC_A)) %>%
  summarise(avg = mean(kfr_pooled_pooled_p25, na.rm = T)) %>%
  as.double()

B <- atlas %>%
  filter(HOLC_B > .5,
         !is.na(HOLC_B)) %>%
  summarise(avg = mean(kfr_pooled_pooled_p25, na.rm = T)) %>%
  as.double()

C <- atlas %>%
  filter(HOLC_C > .5,
         !is.na(HOLC_C)) %>%
  summarise(avg = mean(kfr_pooled_pooled_p25, na.rm = T)) %>%
  as.double()

D <- atlas %>%
  filter(HOLC_D > .5,
         !is.na(HOLC_D)) %>%
  summarise(avg = mean(kfr_pooled_pooled_p25, na.rm = T)) %>%
  as.double()

tibble(HOLC_grade = c("A",
                      "B",
                      "C",
                      "D"),
       avg_kfr_pooled_pooled_p25 = 
         c(44.01, 42.46, 39.87, 36.16)) %>%
  gt()
```

The averages of the kfr_pooled_pooled_p25 for majority A, B, C, and D HOLC
neighborhoods are shown in the table above. It seems that in neighborhoods where
redlining was worse upward mobility is lower than in neighborhoods where
redlining was not prevalent.

\newpage

## Question 8

### Part a

```{r echo=FALSE, warning = FALSE, message=FALSE}

# this looks at the average school district level standardized test scores in
# 3rd grade in 2013

school <- fc %>%
  ggplot(aes(x = gsmn_math_g3_2013, y = kfr_pooled_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. Third Grade Standardized Tests",
       subtitle = "Mobility is higher is areas with higher scores") +
  theme_clean()

# I'm choosing the most recent parent share because that is consistent with the
# school test, but I could have used it from 1990 and then arguably could have
# made more sense

single_parents <- fc %>%
  ggplot(aes(x = singleparent_share2010, y = kfr_pooled_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. Share Single Parent Households",
       subtitle = "Mobility is higher in areas with fewer single parent households") +
  theme_clean()

school
single_parents

```


 - There seems to be a strong relationship between the kfr_pooled_pooled_25
variable and the gsmn_math_g3_2013 variable, which measures the average school
district level sstandardized test scores in third grade in 2013 (where 3 is
where third graders should be, 1 would be where first graders should be, etc.)
It appears that upwarad mobility is higher in areas where students perform
better on these standardized tests.

- There seems to be a strong relationship between the kfr_pooled_pooled_25
variable and the single_parent_share2010 variable, which measures the share of
single-headed households with children 2006-2010. It appears that upward
mobility is higher in areas with a smaller porportion of single parent
households. Note: I could have used the 1990 or 2000 measure for this variable,
but I chose to use the 2010 one so it is consistent with the test score
meassure.

### Part b

```{r echo=FALSE}

# have to drop na values for the cor function to work

# schools correlation

fc_drop <- fc %>%
  select(tract, kfr_pooled_pooled_p25, gsmn_math_g3_2013) %>%
  drop_na()

cor_school <- cor(fc_drop$kfr_pooled_pooled_p25, fc_drop$gsmn_math_g3_2013)

# single parent correlation

fc_drop_sp <- fc %>%
  select(tract, kfr_pooled_pooled_p25, singleparent_share2010) %>%
  drop_na()

cor_sp <- cor(fc_drop_sp$kfr_pooled_pooled_p25, fc_drop_sp$singleparent_share2010)
```

- The correlation of kfr_pooled_pooled_p25 and 
gsmn_math_g3_2013 = `r cor_school`

- The correlation of kfr_pooled_pooled_p25 and 
singleparent_share2010 = `r cor_sp`

## Question 9

### Third Grade Test Scores 

The graphs below show the relationship between upward mobility and 3rd grade
test scores by race. The graphs are binned scatterplots and the correlations
for the relationships can be found in the subtitle of the graphs. Looking at the graphs, it is evident that there are differences in the strength of the correlation between upward mobility and standardized test scores among different racial groups. The correlation between these two variables is strongest for hispanics, followed by whites, asians, and blacks. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# first looking at test scores

school_w <- fc %>%
  ggplot(aes(x = gsmn_math_g3_2013, y = kfr_white_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \n3rd Grade Standardized Tests",
       subtitle = "Race = white, cor = .38") +
  theme_classic()

# new way to drop nas

cor_school_w <- cor(fc$kfr_white_pooled_p25, fc$gsmn_math_g3_2013,
                  use = "complete.obs")
# black

school_b <- fc %>%
  ggplot(aes(x = gsmn_math_g3_2013, y = kfr_black_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \n3rd Grade Standardized Tests",
       subtitle = "Race = black, cor = .29") +
  theme_classic()

cor_school_b <- cor(fc$kfr_black_pooled_p25, fc$gsmn_math_g3_2013,
                  use = "complete.obs")

# hispanic

school_h <- fc %>%
  ggplot(aes(x = gsmn_math_g3_2013, y = kfr_hisp_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \n3rd Grade Standardized Tests",
       subtitle = "Race = hispanic, cor = .52") +
  theme_classic()

cor_school_h <- cor(fc$kfr_hisp_pooled_p25, fc$gsmn_math_g3_2013,
                  use = "complete.obs")

# asian

school_a <- fc %>%
  ggplot(aes(x = gsmn_math_g3_2013, y = kfr_asian_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \n3rd Grade Standardized Tests",
       subtitle = "Race = asian, cor = .32") +
  theme_classic()

cor_school_a <- cor(fc$kfr_asian_pooled_p25, fc$gsmn_math_g3_2013,
                  use = "complete.obs")

# displaying graphs

ggarrange(school_w,school_b,school_h,school_a)
```

### Single Parent Households

The graphs below show the relationship between upward mobility and share of
single parent households by race. The graphs are binned scatterplots and the
correlations for the relationships can be found in the subtitle of the graphs. As with what we saw in the standardized test score analysis, there is variation in the correlation coefficients of upward mobility and single parent household among racial groups. The correlation is strongest for whites and hispanics, followed by asians and blacks. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# single parent share

sp_w <- fc %>%
  ggplot(aes(x = singleparent_share2010, y = kfr_white_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \nSingle Parent Share",
       subtitle = "Race = white, cor = -.45") +
  theme_classic()

# new way to drop nas

cor_p_w <- cor(fc$kfr_white_pooled_p25, fc$singleparent_share2010,
                  use = "complete.obs")
# black

sp_b <- fc %>%
  ggplot(aes(x = singleparent_share2010, y = kfr_black_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \nSingle Parent Share",
       subtitle = "Race = black, cor = -.17") +
  theme_classic()

# new way to drop nas

cor_p_b <- cor(fc$kfr_black_pooled_p25, fc$singleparent_share2010,
                  use = "complete.obs")

# hispanic

sp_h <- fc %>%
  ggplot(aes(x = singleparent_share2010, y = kfr_hisp_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \nSingle Parent Share",
       subtitle = "Race = hispanic, cor = -.46") +
  theme_classic()

# new way to drop nas

cor_p_h <- cor(fc$kfr_hisp_pooled_p25, fc$singleparent_share2010,
                  use = "complete.obs")

# asian

sp_a <- fc %>%
  ggplot(aes(x = singleparent_share2010, y = kfr_asian_pooled_p25)) +
  stat_binmean(n = 20, geom = "point") +
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Upward Mobility vs. \nSingle Parent Share",
       subtitle = "Race = asian, cor = -.24") +
  theme_classic()

# new way to drop nas

cor_p_a <- cor(fc$kfr_asian_pooled_p25, fc$singleparent_share2010,
                  use = "complete.obs")

# displaying graphs

ggarrange(sp_w, sp_b, sp_h, sp_a)
```

