---
title: "empirical_proj_2"
author: "Lindsey Greenhill"
date: "4/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(haven)
library(rdrobust)
library(statar)
library(vtable)
library(ggthemes)
library(sandwich)
library(ggpubr)
library(lmtest)

data <- read_dta("third_grade.dta")
```

## Question 1

The fundamental problem of causal inference is that you cannot see the outcome
for a child who both was in a small class and not in a small class. In other
words, it is impossible to measure the dual outcomes when a child both is in a
small class and not in a small class, because the child can only either be in a
small class or not in a small class.

## Question 2

The following variables have missing values:

- towncode: 318 missing
- math: 58 missing
- verb: 466 missing
- ses_index: 318 missing
- boy: 1037 missing

```{r echo=FALSE}
summary(data)
```

## Questions 3 and 4

```{r echo=FALSE, results = "asis"}

# getting the means

math_mean <- mean(data$math, na.rm = T)
verb_mean <- mean(data$verb, na.rm = T)

# getting the sds

math_sd <- sd(data$math, na.rm = T)
verb_sd <- sd(data$verb, na.rm = T)

# standardizing

data$math_std <- (data$math - math_mean) / math_sd
data$verbal_std <- (data$verb - verb_mean) / verb_sd

# creating test score index

data <- data %>%
  mutate(test_score_index = .5 * (math_std + verbal_std)) %>%
  mutate(math_miss = if_else(is.na(math),
                             1,
                             0),
         verb_miss = if_else(is.na(verb),
                                1,
                                0)) %>%
  mutate(test_score_index = if_else(is.na(test_score_index),
                                    if_else(math_miss == 1, verbal_std, math_std),
                                    test_score_index))


# summary table

sumtable(data,
         vars = c("math", "verb", "test_score_index"),
         summ = c("mean(x)",
                  "sd(x)",
                  "min(x)",
                  "max(x)"),
         out = "latex")

```

## Question 5

All of the histograms appear to not by symmetrical and all of them have a long
left tail.

```{r echo=FALSE, warning = FALSE}

# math hist

hist_math <- data %>%
  ggplot(aes(x = math,
             y = ..density..)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Math Scores") +
  theme_clean()


# verbal hist

hist_verb <- data %>%
  ggplot(aes(x = verb,
             y = ..density..)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Verbal Scores") +
  theme_clean()

# index hist

hist_index <- data %>%
  ggplot(aes(x = test_score_index,
             y = ..density..)) +
  geom_histogram(binwidth = .25) +
  labs(title = "Distribution of Test Score Index") +
  theme_clean()

ggarrange(hist_math,
          hist_verb,
          hist_index)

```

## Question 6

```{r echo=FALSE, warning = FALSE}

# creating binned scatter plots

bin_index_ses <- data %>%
  ggplot(aes(x = ses_index, y = test_score_index)) +
  stat_binmean(n = 20, geom = "point") +
  labs(title = "Test Score Index vs. SES Index") +
  theme_classic()

bin_class_ses <- data %>%
  ggplot(aes(x = ses_index, y = class_size)) +
  stat_binmean(n = 20, geom = "point") +
  labs(title = "Class Size vs. SES Index") +
  theme_classic()
  
bin_index_class <- data %>%
  ggplot(aes(x = class_size, y = test_score_index)) +
  stat_binmean(n = 20, geom = "point") +
  labs(title = "Test Score Index vs. Class Size") +
  theme_classic()

ggarrange(bin_index_ses,
          bin_class_ses,
          bin_index_class)

```

## Question 7

As we can see from the binned scatter plots above, the socio economic
index variable is correlated with both class size and test score index.
As such, it is a counfounding variable and mean we cannot interpret 
the relationship between class size and test scores causally because
it is likely that a student's socioeconomic status influences both 
outcomes, meaning that we don't know how greatly class size affects
test scores versus ses affects test scores.

## Question 8

### Part a

I chose a quadratic model it seems to fit the data better than a linear model,
especiaally for the data to the left of the break

```{r echo=FALSE}

narrow <- data %>%
  filter(school_enrollment <= 80)

# I chose a quadratic because it seems to fit the data better, especially for
# the data to the left of the break

rdplot(narrow$class_size,
       narrow$school_enrollment,
       c = 41,
       p = 2,
       nbins = c(20,20),
       y.lim = c(0, 40),
       binselect = "es",
       title = "Class size vs. total school enrollment",
       x.label = "Total School Enrollment in 3rd Grade",
       y.label = "Class Size")
```

### Part b

I chose to use a linear model because it seems to fit the data better than a
quadrtatic model. When I chose a quadrataic model, it seemed like residuals on
the right side of the break were much more in the positive than the negative,
which would be bad. Looking at the graph, it doesn't seem like there is a wide
gap in schools with 40 kids and schools with 41 kids enrolled. 

```{r echo=FALSE}

# switching the varibale from above
rdplot(narrow$test_score_index,
       narrow$school_enrollment,
       c = 41,
       p = 1,
       nbins = c(20,20),
       y.lim = c(-.75, .5),
       binselect = "es",
       title = "Test Score Index vs. Total School Enrollment",
       x.label = "Total School Enrollment in 3rd Grade",
       y.label = "Test Score Index")
```

## Question 9

### Part a

```{r echo=FALSE}
# religious, boy (with test score index), born_isr

by_school <- data %>%
  group_by(schlcode) %>%
  mutate(avg_boy = mean(boy, na.rm = T),
         avg_religious = mean(religious, na.rm = T),
         avg_born_isr = mean(born_isr, na.rm = T)) %>%
  filter(school_enrollment <= 80)

# there doesn't appear to be a significant gap in % boys

rdplot(by_school$avg_boy,
       by_school$school_enrollment,
       c = 41,
       p = 1,
       nbins = c(20,20),
       y.lim = c(0, 1),
       binselect = "es",
       title = "% Boys vs. Total School Enrollment",
       x.label = "Total School Enrollment in 3rd Grade",
       y.label = "% Boys")

# there doesn't appear to be a significant gap in % borrn in isr

rdplot(by_school$avg_born_isr,
       by_school$school_enrollment,
       c = 41,
       p = 1,
       nbins = c(20,20),
       y.lim = c(.5, 1.2),
       binselect = "es",
       title = "% Born in Isr vs. Total School Enrollment",
       x.label = "Total School Enrollment in 3rd Grade",
       y.label = "% Born in Isr")

```

### Part b

The identification assumption for the regression discontinuity design is 
that schools on either side of the cutoff are similar in all ways and as such
the fact that they have school enrollment over or under the cutoff is as if
random. The graphs from part a above are consistent with this identification
assumption, as schools just at either side of the cutoff do not seem to have
different percentages of boys or kids born in Israel enrolled.






